```{r}
library(arules)
library(arulesViz)
library("RColorBrewer")
str(df)
```

(2 points)Read the file groceries which contains data about transactions made in the market.
Make the data usable for Association Rules Mining. 
(Hint. You can use function read.transactions)
plot the data to see what are the first 5 most frequent items
```{r}
txn <- read.transactions("groceries.csv",rm.duplicates = FALSE, cols = NULL, sep = ",")
itemFrequencyPlot(txn,
   topN=10,
   col=brewer.pal(8,'Pastel2'),
   main='Relative Item Frequency Plot') 
```
Most frequent five items are whole milk, other vegetables, rolls/buns, soda and yogurt.

(5 points) Using appriori algorithm, create the rules by setting the support to be 0.005 and confidence to be 0.1 Set the length of the rules to be 4.
Interpret the meaning of support and confidence.
```{r}
rules <- apriori(txn, parameter = list(minlen = 4, maxlen = 4, supp = 0.005, conf = 0.1))
```
Support for itemset is the percentage of transactions which include the itemset.
Confidence is the percent of antecedent transactions which also contain the consequent itemsets.
We only allow the itemset with above the threshold values of 0.005 and 0.001 to be considered.

(7 points) Inspect the first 5 rules. Sort them by support in descending order. 
Interpret the first 2 rules. What is the meaning of support, comfidence and lift?
```{r}
inspect(rules[1:5])
rules_df <- as(rules, "data.frame")
rules_df[order(-rules_df$support),]
```
1  {fruit/vegetable juice, other vegetables, yogurt} => {whole milk} 
Confidence shows that there is 0.61 probability that we will have whole milk bought given the customer bought the items on the left.
Lift ratio shows how effective is the rule finding consequents (confidence / the overall probabilty of the right side).
The more frequent is the itemset on the right, the lower is the lift ratio and the higher is the confidence, the higher the lift.
Support shows overall impact. The probability that the itemset will occur is 0.005.

2 {fruit/vegetable juice, whole milk, yogurt} => {other vegetables} 
Confidence shows that there is 0.53 probability that we will have the vegetables and "other vegetables" bought given the customer bought the items on the left.
Lift ratio shows how effective is the rule finding consequents (confidence / the overall probabilty of the right side).
The more frequent is the itemset on the right, the lower is the lift ratio.
Support shows overall impact. The probability that the itemset will occur is 0.005.


(6 points) Subset the rules to the ones that contaiin only "whole milk" in the right hand side and with confidence more than 60%. 
What group of products has the highest support?
```{r}
milk_rules <- subset(rules, rhs %ain% 'whole milk')
milk_rules <- subset(rules, confidence > 0.6)
milk_rules
inspect(milk_rules[1:5])
```
The ones which are more frequent, (higher count) have higher support.

(10 points) Explore the library(arulesViz) and make 3 insightful vizualisations on the data. 
```{r}
plot(milk_rules, type = "scatterplot")
plot(milk_rules, method = "graph")
plot(milk_rules, method = "grouped")
```